Arms Transfers and Frozen Conflict Dynamics

### 2024-04-03 paper.qmd
Notes

@@

@@

such that $\mathbb{E}\left[y_{t} | S_t = k\right] = \mu_{ik}$ and $\text{Var}(y_{t} | S_t = k) = \mu_{ik} + \frac{\mu_{ik}^2}{\phi}$.

@@

@@

@@

Future work:
- incorporate more battlefield dynamics
- dyadic level rather than conflict level

@@

>>>>

Central research Question:

How do international arms transfers to incumbent governments affect the dynamics of civil conflict?
    - How do international arms transfers affect the probability of civil conflicts transitioning between frozen & active

one-sided international arms transfers

@@

TODO: 3 state model, no restriction on transitions (ie 1 -> 3 and 1 -> 2)

TODO: hidden markov models
  - Allows us to estimate the probability that a conflict is in a particular state at a given time point (ie estimate latent states)
  - Allows us to model the effects of covariates on the probability of transitioning between latent states & model the effects of covariates on observable outcomes

Note: "a hidden markov model is a tool for representing probability distributions over sequences of observations"

Potential issues: computational costs, and multimodality in the posterior distribution, driven by the high variance in observed fatalities and the difficulty in identifying latent states.
    - One possible mitigation, informative priors. It pays to be bayesian. (Maybe compare bayesian approach to only EM algorithm).

@@

such that $\mathbb{E}\left[y_{t} | S_t = k\right] = \mu_{ik}$ and $\text{Var}(y_{t} | S_t = k) = \mu_{ik} + \frac{\mu_{ik}^2}{\phi}$.[^negbin]

[^negbin]: This parameterization arises naturally from ... poisson-gamma mixture. As $\phi \to \infty$ then $y | S_t, \theta \overset{d}{\to} \text{Pois}$.

@@

\begin{equation}
p(\theta | \boldsymbol{y}, X) \propto \prod_{i=1}^N p(y^{(i)}_{1:T_i} | X^{(i)}, \theta) p(\theta)
\end{equation}

@@

Because of non-linear model, empirical estimands

average posterior predicted probability for transitioning from state $j$ to state $k$ for some constant TIV value across all transition cases.

Formally,

\[ g_{j \to k}(\theta, \boldsymbol{X}^\star) = \frac{1}{N} \sum_{i=1}^N \frac{1}{T_i - 1} \sum_{t=2}^{T_i} P(S_t^{(i)} = k | S_{t-1}^{(i)} = j, X_t^{\star(i)}, \theta) \]

where $\boldsymbol{X}^{\star(i)} = [c, X_2^{(i)}, \ldots, X_D^{(i)} ]$ for $c \in [0, \max_{i \in \{1, \ldots, N \}} X_1^{(i)}]$

For a fixed $\boldsymbol{X}^\star$, the push-forward distribution is approximated via monte carlo integration with draws from the posterior.

\[ p(g_{j \to k}(\theta, \boldsymbol{X}^\star) | \boldsymbol{y}, \boldsymbol{X}) \approx \{ g_{j \to k}(\theta^{(m)}, \boldsymbol{X}^\star) : \theta^{(m)} \sim p(\theta | \boldsymbol{y}, \boldsymbol{X}) \} \]

Plots:
1. Predicted vs. actual deaths
2. Predicted probability of not being frozen
3. Parameter plot for beta tiv
4. APP - Marginal Effects
    - Prob. of a frozen conflict escalating
    - Prob. of an actie conflict de-escalating
5. (Maybe?) long-run effect of tiv (low vs medium vs high w/ covariates at means)

@@

Shortcomings:
- Frozen conflict measure does not account for the ability of actors
  to re-escalate a conflict --- no distinction between a weakened
  insurgency versus a de facto separatist state
- Excluded low intensity conflicts

Future work:
- incorporate more battlefield dynamics, including distance to pop. centers, proportion of calendar year with violent events etc etc.
- dyadic level rather than conflict level

@@

# Appendix

1. Forward-backward algorithm (argue for why forward-backward instead of viterbi)
2. Simulation based calibration
3. ppc_dens_overlay plots
4. Table of \beta & \lambda estimates for all independent variables
5. Robustness checks
